\textbf{Betelwise1: A Hybrid Conventional-CNN Approach for Efficient Planetary Seismic Event Detection and Analysis}

### 2025-05-24 paper.tex
update content

    \subsection{ Martian Seismology}
    Mars has been the focus of renewed seismic interest. The first attempt to place a seismometer on Mars was with the
    Viking landers in 1976. The Viking 2 lander's seismometer, mounted on the lander deck, provided data primarily on
    wind-induced lander vibrations, and only one candidate seismic event was tentatively identified (Anderson et al.,
    1977; Lognonné, 2005). The on-deck deployment and instrument sensitivity were major limitations.

@@

    % \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73. 

@@

    % \bibitem{b10} W. B. Banerdt et al., ``Initial results from the InSight mission on Mars,'' \emph{Nature Geoscience}, vol. 13, pp. 183--189, Mar. 2020.

>>>>

    \subsection{ Martian Seismology}
        Mars has been the focus of renewed seismic interest. The first attempt to place a seismometer on Mars was with
        the Viking landers in 1976. The Viking 2 lander's seismometer, mounted on the lander deck, provided data
        primarily on wind-induced lander vibrations, and only one candidate seismic event was tentatively identified
        \cite{Anderson1977,Lognonne2005}. The on-deck deployment and instrument sensitivity were major limitations.

        Decades later, the NASA InSight mission, which landed in 2018, successfully deployed the SEIS (Seismic
        Experiment for Interior Structure) instrument directly onto the Martian surface \cite{Lognonne2019, Banerdt2020}. 
        %(Lognonné et al., 2019; Banerdt et al., 2020) 
        SEIS comprises a suite of highly sensitive three-component Very Broad Band (VBB) and three-component Short
        Period (SP) seismometers. InSight has definitively recorded hundreds of marsquakes, confirming Mars as a
        seismically active planet \cite{Giardini2020,Clinton2021}. Martian seismic signals, like lunar ones, can be
        influenced by near-surface scattering, though generally less intense than on the Moon. The Marsquake Service
        (MQS) is responsible for ongoing event detection and cataloging, employing a combination of traditional and
        evolving ML-assisted techniques \cite{Clinton2021}.

        %TODO add a section on only CNNs ================================================================

    \subsection{ Machine Learning Paradigms in Seismology }
        The increasing volume and complexity of seismic data from both terrestrial and planetary missions have spurred the
        adoption of machine learning techniques across various seismological tasks. Deep learning, in particular CNNs and
        Recurrent Neural Networks (RNNs), has shown remarkable success in earthquake detection, phase picking, seismic
        imaging, and source characterization on Earth \cite{MousaviBeroza2022}.
        %(Mousavi and Beroza, 2022; Mousavi et al., 2020 - "Earthquake Transformer").

        CNNs are adept at automatically learning hierarchical features from data, making them suitable for analyzing raw
        seismic waveforms or their time-frequency representations (e.g., spectrograms) (Perol et al., 2018; Ross et al.,
        2018b). Generative Adversarial Networks (GANs) have also been explored, for instance, by Li et al. (2018) who
        used a GAN critic as an automatic feature extractor combined with a Random Forest classifier for discriminating
        earthquake P-waves from noise in early warning systems. This highlights the potential of ML to not only classify
        but also to learn compact and effective representations of seismic waves.

\section{Framework}
    %Overview
    Overview
    \subsection{Datasets and Input Seismogram Preparation}
    hello
        \subsubsection{Data Sources (Lunar and Martian)}
            hello

@@

    \bibitem{Anderson1977} D. L. Anderson, W. F. Miller, G. V. Latham, Y. Nakamura, M. N. Toksöz, A. M. Dainty, F. K. Duennebier, A. R. Lazarewicz, R. L. Kovach, and T. C. D. Knight, ``Seismology on Mars,'' \emph{J. Geophys. Res.}, vol. 82, no. 28, pp. 4524--4546, 1977.
    \bibitem{Banerdt2020} W. B. Banerdt et al., ``Initial results from the InSight mission on Mars,'' \emph{Nature Geoscience}, vol. 13, pp. 183--189, Mar. 2020.
    \bibitem{Giardini2020} D. Giardini et al., ``The seismicity of Mars,'' \emph{Nature Geoscience}, vol. 13, pp. 205--212, 2020.
    \bibitem{Clinton2021} J. F. Clinton et al., ``The Marsquake catalogue from InSight, sols 0--478,'' \emph{Phys. Earth Planet. Inter.}, vol. 310, Art. no. 106595, 2021
    \bibitem{MousaviBeroza2022} S. M. Mousavi and G. C. Beroza, ``Deep-learning seismology,'' \emph{Science}, vol. 377, no. 6606, p. eabm4470, Aug. 2022, doi: 10.1126/science.abm4470.   % \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73. 

@@
